Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-05-03 doc/RMD/mlg-mcg.Rmd
update to inclue top 5 graph

Now we can take a look at the concordance of MLGs to MCGs

```{r}
mll.custom(dat11) <- strata(dat11)$MCG
mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  as_data_frame() %>%
  filter(Freq > 0)
mcgs <- mcgmlg %>%
  group_by(MCG) %>%
  summarize(MLGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
  arrange(desc(MLGs))
mlgs <- mcgmlg %>%
  group_by(MLG) %>%
  summarize(MCGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
  arrange(desc(Samples), desc(MCGs))
mcgs
mlgs
any(mcgs$MLGs > 1)
```

@@

```{r}
mlg.filter(dat11, dist = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps + (0.5/11)
dat11

@@

```{r}
gdf <- mcgmlg %>% 
  mutate(MLG = paste0('MLG.', MLG))
MLGS <- gdf %>% 
  group_by(MLG) %>%
  summarize(size = sum(Freq)) %>%
  rename(vertex = MLG)
MCGS <- gdf %>% 
  group_by(MCG) %>%
  summarize(size = sum(Freq)) %>%
  rename(vertex = MCG)
VAT <- bind_rows(MLGS, MCGS)
g <- gdf %>% 
  select(MCG, MLG) %>%
  graph_from_data_frame(vertices = VAT)
osize <- V(g)$size
```

@@

V(g)$size <- sqrt(osize)/10
V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
set.seed(2017-05-03)
lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")

@@

mcg_mlg_graph
ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
       width = 88*3, height = 88*3, units = "mm")
```

>>>>

Now we can take a look at the concordance of MLGs to MCGs. We can do this by 
creating a contigency table. Of course, since we have 87 and well over 100 MLGs,
this means that the contingency table is going to be big, so to summarize it
further, I'm creating two tables, one based on MLGs that will count the number
of MCGs within each MLG and vice-versa. Of course we lose information like, if
an MCG contains several MLGs, how can we tell what the abundance is? A handy 
measure is Evenness, which scales from 0 to 1, indicating how skewed the 
observations are.

```{r}
mll.custom(dat11) <- strata(dat11)$MCG
mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  as_data_frame() %>%
  filter(Freq > 0)
mcgs <- mcgmlg %>%
  group_by(MCG) %>%
  summarize(MLGs = sum(Freq > 0), 
            Samples = sum(Freq), 
            Evenness = diversity_stats(Freq)["E.5"], 
            data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
  arrange(desc(MLGs))
mlgs <- mcgmlg %>%
  group_by(MLG) %>%
  summarize(MCGs = sum(Freq > 0), 
            Samples = sum(Freq), 
            Evenness = diversity_stats(Freq)["E.5"], 
            data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
  arrange(desc(Samples), desc(MCGs))
mcgs
mlgs

```

It might be better to visualize these data as barplots. Here we are mapping the
type (MCG/Count) to color and the opacity (alpha) to Evenness.

```{r barplots, fig.width = 10, fig.height = 4}
mcgs %>% 
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = type, fill = type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  annotate(geom = "text", x = 13, y = 51, label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
  ggtitle("How Evenly are Multilocus Genotypes (MLGs) spread across MCGs?")
mlgs %>% 
  gather(type, count, MCGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
  ggplot(aes(x = MLG, y = count, group = type, fill = type)) +
  geom_col(aes(width = ifelse(type == "MCGs", 0.5, 0.85), alpha = Evenness), color = "black", position = "identity") +
  annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
  ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
```

@@

```{r}
mlg.filter(dat11, dist = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps + (3*(0.5/11))
dat11

@@

```{r}
make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MCG, MLG) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}
g <- make_mcgmlg_graph(mcgmlg)
gf <- make_mcgmlg_graph(fmcgmlg)
osize <- V(g)$size
fosize <- V(gf)$size
```

@@

V(g)$size <- sqrt(osize)/10
V(gf)$size <- sqrt(fosize)/10
set.seed(2017-05-03)
lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")
flay2 <- create_layout(gf, layout = "igraph", algorithm = "nicely")

@@

mcg_mlg_graph
mcg_mlg_graph %+% flay2
ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
       width = 88*3, height = 88*3, units = "mm")
```

## Subgraph

So that's a big hairy graph. What happens when we look at a subgraph of the
top 5 MCGs?

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
set.seed(2017-05-03)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
```

Vey nice!
